Fail Faster

### 2025-03-13 paper/paper.tex

\jwc{Same story in scala. Approximately 2x difference for the same generator... (scala uses a slightly different generator type, but the principles are the same)}

@@

@@

\jwc{
The problem is plain: while the abstractions that PBT libraries like base quickcheck provide are indespensible for writing idiomatic generators, the performance overhead of using them is dramatic.
The compiler (and flambda) uses heuristics (\href{https://ocaml.org/manual/5.3/flambda.html}{here}) to decide when to specialize and inline code, and these heuristics cannot guarantee that generator abstractions are zero-cost.
These heuristics are also extremely conservative about inlining and specializing recursive functions, which all generators of recursive data types must be.
}

>>>>

@@

\jwc{Same story in scala. 
Approximately 2x difference for the same generator (though an order of magnitude slower overall than bq)... (scala uses a slightly different generator type, but the principles are the same)
This doesn't even account for more inlining we could do to eliminate the boxing/unboxing of results. (Scalacheck generators can fail, though this adds significant overhead)
}

@@

\jwc{
The problem is plain: while the abstractions that PBT libraries like base quickcheck provide are indespensible for writing idiomatic generators, the performance overhead of using them is dramatic.
Modern compilers use heuristics to decide when to specialize and inline code \cn, and these heuristics cannot guarantee that generator abstractions are zero-cost.
These heuristics are also necessarily conservative about inlining and specializing recursive functions, which all generators of recursive data types must be.
}
